---
title: "Place effects"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
master_data <- read_csv("all_segs.csv") %>%
  mutate(Final = Position == "Final",
         LongV = Long.Short == "long",
         Duration = Duration * 1000,
         Manner = case_when(
           Seg %in% c("n", "l") ~ "Sonorant",
           Seg %in% c("s", "ss") ~ "Fricative",
           Seg == "m" ~ "Labial nasal",
           TRUE ~ "Stop"),
         Voiceless = Seg %in% c("p", "t", "k", "s", "ss"),
         id = paste(File, Seg, Duration_in_ms, sep = "")) %>%
  select(-Duration_in_ms, -Long.Short, -Position)

sonorants <- master_data %>%
  filter(Manner == "Sonorant") %>%
  mutate(Fortis = str_detect(Word, "^[^aeiouáéíóú]*[aeiouáéíóú]+(nn|ll)"))

master_data <- master_data %>%
  left_join(sonorants)
```

```{r}
stops_data <- master_data %>%
  filter(Manner == "Stop")

fit0 <- lm(Duration ~ Seg + Final, data = stops_data)
summary(fit0)
```
#so, as you'd expect: with /b/ as the starting point, we see no sig. dif. for the other voiced stops: /d/ and /g/, and large increases in duration for the voiceless stops, /k/ , /p/, /t/, as well as a significant finality effect



```{r}
fit1 <- lm(Duration ~ Seg * Final, data = stops_data)
summary(fit1)
```
#in this model, the interaction between each segment and finality is examined: we see similar but slightly reduced increased duration for the voiceless stops. Finality no longer shows a sig. effect (p ~ 0.4). The only sig. interaction effect seems to be for /k/ and final postion (112.289, p < 0.001)

```{r}
fit2 <- lm(Duration ~ Seg * Final + LongV, data = stops_data)
summary(fit2)
```
#Now, with the addition of preceding long vowel, we see no sig effect (LongVTRUE = -8.462, p > 0.5)

```{r}
fit3 <- lm(Duration ~ Seg * Final + LongV + LongV:Final, data = stops_data)
summary(fit3)
```
#This model seems promising, as we see the same lengthening for all three voiceless stops, none for voiced stops. 
#Also, no sig. effect for finality alone, or for long vowel, but again a sig. interction between /k/ and finality.
#with the addition of the interaction of finality and preceding long vowel, we see a marginally sig.effect (FinalTRUE:LongVTRUE: -23.2262, p ~ 0.035) 

```{r}
fit4 <- lm(Duration ~  Seg * Final + LongV + LongV:Final + LongV:Seg, data = stops_data)
summary(fit4)
```
```{r}
fit5 <- lm(Duration ~ Seg * Final * LongV, data = stops_data)
summary(fit5)
```
```{r}
AICc(fit0, fit1, fit2, fit3, fit4, fit5)
```
fit3 seems to be the best model

```{r}
stops_short <- filter(stops_data, LongV == FALSE)
stops_short_marked <- mutate(stops_short, ShortV = TRUE)
stops_long <- filter(stops_data, LongV == TRUE)
stops_long_marked <- mutate(stops_long, ShortV = FALSE)
stops_short_column <- rbind(stops_short_marked, stops_long_marked)
```

```{r}
fit6 <- lm(Duration ~ Seg * Final + ShortV + ShortV:Final, data = stops_short_column)
summary(fit6)
```
```{r}
AICc(fit6)
```
```{r}
fit7 <- lm(Duration ~ Seg * Final * ShortV, data = stops_short_column)
summary(fit7)
```
```{r}
AICc(fit7)
```
#this model was an attempt to look at the interaction of each segment and final position and a preceding short vowel individually. The model seems to have become quiet convoluted though. the only extra sig. effect is of d and short vowel, which is actually shorter in duration (-44, p ~ 0.015)
