---
title: "sons regression models"
output: html_notebook
---
```{r}
library(tidyverse)
```

```{r}
sons_data <- master_data %>%
  filter(Manner == "Sonorant")

fit0 <- lm(Duration ~ Fortis + Final, data = sons_data)

summary(fit0)
```
Fortis gives 31 longer 
Final gives 15 longer
Both have low p numbers so seem reliable

```{r}
fit1 <- lm(Duration ~ Fortis * Final, data = sons_data)

summary(fit1)
```
interesting changes when interaction of fortis and finality considered:

the effect of fortis is reduced (20 rather than 31), and effect of finality profoundly reduced (7 rather than 15, with a high p number: ~0.03) 

the interaction fortis * final seems to have a large effect (25, with a low p-number)
```{r}
fit2 <- lm(Duration ~ Fortis * Final + LongV, data = sons_data)
summary(fit2)
```
when the preceding vowel length is taken into account (additive model) there is a simiar effect for fortis (19, with a low p-number), and for fortis (8 with a high p-number). The vowel does not seem to have an effect in this model (p number > 0.2).

Again the interaction of fortis*final has a significant effect (24 with a low p-number)

```{r}
fit3 <- lm(Duration ~ Fortis * Final + LongV + LongV:Final, data = sons_data)
summary(fit3)
```
This model reiterates that fortis has a signifcant lengthening effect (19 with a low p-number) but final, in this model, does not (5 with a high p-number, > 0.2). preceding vowel length, as expected, has no significant effect (small negative, as in fit 2, with p > 0.1). 

Again, the interaction of final and fortis has a significant effect: 27 with a low p-number



```{r}
fit4 <- lm(Duration ~ Fortis * Final + LongV + LongV:Final + LongV:Fortis, data = sons_data)
summary(fit4)
```
confirmed from previous models:

effect of fortis (19 low p-number)
no effect of final (5, p > 0.2)
no effect of long (-6, p > 0.1)
effect of fortis*final interaction (27, low p number)

New:

no effect of final*long interaction (~7, p ~ 0.3)
no effect of fortis*long interaction (0.2, p ~ 1 )

```{r}
fit5 <- lm(Duration ~ Fortis * Final * LongV, data = sons_data)
summary(fit5)

```
confirmed: 

fortis effect (19, low p)
no final effect (5, p ~ 0.2)
no preceding vowel effect (-5, p ~ 0.2)
fortis*final interaction effect (25, low p)
no effect of fortis*long interaction (-3, p ~ 0.75 )
no effect of final*long interaction (~5.5, p ~ 0.4)

New:

no effect of fortis*final*long (12, p ~ 0.5)


```{r}
AICc(fit0, fit1, fit2, fit3, fit4, fit5)
```
fit1 seems to be the 'best' model according to the AICc number


```{r}
random_fit <- lmer(Duration ~ Fortis * Final + LongV + (1|Word), data = sons_data)

summary(random_fit)

```
```{r}
random_fit0 <- lmer(Duration ~ Fortis * Final + (1|Word), data = sons_data)
summary(random_fit0)
```
```{r}
random_fit1 <- lmer(Duration ~ Fortis * Final + LongV + LongV:Final + (1|Word), data = sons_data)
summary(random_fit1)
```


```{r}
anova(random_fit, random_fit0, random_fit1, test = "Chisq")
```

